Software Clusterings with Vector Semantics and the Call Graph

### 2019-04-27 manuscript/manuscript.tex

%
% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
% other information printed in the page headers. This command allows the author to define a more concise list
% of authors' names for this purpose.
\renewcommand{\shortauthors}{Papachristou}

@@

\noindent \textbf{Method -- } Cluster files by combining document embeddings, generated with the  \emph{Doc2Vec} algorithm, and the call graph, provided by Static Graph Analyzers to an augmented graph.
Use of the \emph{Louvain Algorithm} to iteratively exhibit its community structure and propose a module--level clustering.

@@

\noindent \textbf{Conclusions --} Semantic information from vector semantics and the call graph can produce 
accurate results for software clusterings of large systems.

@@

This paper aims to propose software clusterings using a combination of vector semantics on the source code provided 
by \emph{Doc2Vec} as well as the existing structure of 
the codebase via the call graph in order to form communities and arrive at a module-level 
architecture viewed from the call graph perspective. 
We will study Linux codebase and we will compare it with ground truth in using the MoJo metric \cite{mojo} as well as 
compare it with other state-of-the-art and baseline clustering algorithms \cite{maqbool_overview, evaluation} in terms of 
\emph{stability}, \emph{authoritativeness} and \emph{extremity}

@@

For the prepossessing stage use the NLP package spaCy \cite{spacy} which provided us with very useful tools for the extraction. 
After prepossessing, we use Gensim \cite{gensim} to train the Doc2Vec model. 
First of all, we tokenize the code and remove the stop-words. 
After that, for each token, we split it into its constituent parts using dynamic programming and lemmatize 
each individual sub-token. 
For example the method \texttt{\_\_zone\_seqlock\_init} corresponds to \texttt{zone}, \texttt{seqlock} and \texttt{init}, 
\texttt{inprogress} is split into \texttt{in} and \texttt{progress} and \texttt{literals} becomes \texttt{literal}.

\paragraph{Call Graph Generation} 
The codebase is processed by a Static Graph Analyzer. 
In our approach, focusing on C projects, we use \emph{CScout} \cite{cscout} in order to generate the directed call graph. 
Each source code file is assigned to a module. 
Modules can be user defined or automatically generated by their respective directories at a desired directory tree depth.

@@

\paragraph{Experimental Setting} Our method was run on Linux 4.21,
consisting of 20.3 million SLOC against Average-Linkage \cite{average}, Complete-Linkage \cite{complete} 
and Ward-Linkage \cite{ward} as well as ACDC and LIMBO with $B = 100$, $S = \infty$ and Bag-of-Words features. The results are presented in Table \ref{tab:evaluation}. 
As ground truth we have used the first level directories as a target clustering and as input we have considered 
the modules of the one-top directories. 
For example, the source code file \texttt{drivers/net/ieee802154\-/mcr20a.c} has a ground truth value of \texttt{drivers} 
and it is considered under the same module as every \texttt{.c} and \texttt{.h} file under \texttt{drivers/net/ieee802154}. 
As a clustering distance metric we have used the MoJo metric, as proposed in reference \cite{mojo}. 
Since, Louvain Community Detection produces  different (but very similar) results every time it runs, we ran the simulations 
multiple times and averaged the results.

@@

In Table \ref{tab:evaluation}, there are the results of the clusterings ordered in increasing MoJo distance 
with respect to the Ground Truth. 
We have included the number of clusters produced, the cluster size range, the average cluster size the standard
deviation in the cluster sizes, the median cluster and the MoJo distances with respect to the ground truth. 
The Ground Truth clustering is also present in the table. 
Our quality metrics for the clusterings are 
\emph{extremity}, \emph{authoritativeness} and \emph{stability} as proposed in reference \cite{evaluation}

Firstly, our approach produced balanced clusterings thus demonstrating low \emph{extremity} compared to the other clusterings, 
producing an amount of clusters that was close to the Ground Truth with smaller standard deviation in the cluster sizes than 
the other clustering methods, without knowing the number of clusters \emph{a priori}. 
Besides this, our approach also seems to give balanced clusters with respect to \emph{Median Cluster} criterion. 
Note that ACDC, which took into account only the structural properties of the files, produced a high 
granularity clustering that failed to catch the structure of Linux Codebase. 
Moreover, in the \emph{authoritativeness} criterion, our approach outperformed the other clusterings to a very up to 10 times 
in terms of MoJo distance being very close (smaller is better) to the Ground Truth, with a very small standard deviation. 
This also suggests great \emph{stability} in the resulting clusterings. 
Note that the other clustering methods we compared against ours gave disappointing MoJo distances even with their cluster
number defined equal to the one of the Ground Truth.

@@

Onaiza Maqbool and Haroon A. Babri \cite{maqbool_overview} present an
overview of the various approaches taken towards to hierarchical clustering algorithms
for software systems. In their study, they compare the various clustering algorithms
such as ACDC, LIMBO and other Traditional Clustering Algorithms such as Single
Linkage, Complete Linkage \cite{complete} and Weighted Linkage presenting their efficiency regarding 
multiple feature sets. 
Added to this, a \emph{semantics-based architectural view} of the system, as discussed in reference
\cite{large_study} reveals significant aspects of a software system and its change over 
time, which suggests that semantic-based approached should be followed for gaining better 
understandings of a software system. Their suggestion is a \emph{term frequency--inverse 
document frequency} (TF-IDF) model serving as a similarity metric between source code 
entities. 
Work in Software Clustering Algorithms (SCA) has been extensively done by Tzerpos in
references \cite{mojo}, \cite{stability}, \cite{acdc} and \cite{limbo} with the ACDC
and LIMBO clustering algorithms.

In particular, the ACDC algorithm leans toward to software components comprehension 
based on subsystem patterns. Their approach considers an initial structure of the system, 
without taking into account semantics, and tries to build comprehensive clusterings of the given ground structure. 
The authors also conducted a study on an older version of  Linux. 
Moreover, Andritsos and Tzerpos in reference \cite{limbo} present an Information--Theoretic 
approach of SCA by developing LIMBO which clusters modules upon inserting their Distributional Cluster Features 
to a B+-tree variant and then applying the \emph{Agglomerative Information Bottleneck} algorithm, merging the 
clusters with the minimum mutual information loss together.

\section{Conclusions} This paper is set out to show how we can use a combination of vector semantics
and information from the call graph in order to produce meaningful clusterings of a software system.
We conclude that we have outperformed state-of-the-art software clusterings and conventional
agglomerative clustering algorithms 
in terms of \emph{extremity}, \emph{authoritativeness} and \emph{stability} 
without knowing the number of clusters of the Ground Truth. This evidence strongly
supports the further usage of vector semantics and the call graph for architecture recovery.
Finally, further work is encouraged in several areas such as 
integration with more static analyzers and use of the project in order to propose layered architectures.

>>>>

\renewcommand{\shortauthors}{Papachristou}

@@

\noindent \textbf{Method -- } Cluster files by combining document embeddings, generated with the  \emph{Doc2Vec} algorithm, and the \emph{call graph}, provided by Static Graph Analyzers to an augmented graph.
Use of the \emph{Louvain Algorithm} to iteratively exhibit its community structure and propose a module--level clustering.

@@

\noindent \textbf{Conclusions --} Semantic information from vector semantics as well as the call graph can produce 
accurate results for software clusterings of large systems.

@@

This paper aims to propose software clusterings using a combination of vector semantics on the source code provided 
by \emph{Doc2Vec} as well as the existing structure of 
the codebase via the call graph in order to form communities and arrive at a module-level 
architecture viewed from the call graph perspective. 
We will study the Linux codebase and we will compare it with ground truth in using the MoJo measure \cite{mojo} as well as 
compare it with other state-of-the-art and baseline clustering algorithms  in terms of \emph{stability}, \emph{authoritativeness} and \emph{extremity}
\cite{maqbool_overview, evaluation}.

@@

For the prepossessing stage use the NLP package spaCy \cite{spacy} which provided us with very useful tools for the extraction. 
After prepossessing, we use Gensim \cite{gensim} to train the Doc2Vec model. 
First of all, we tokenize the code and remove the stop-words. 
After that, for each token, we split it into its constituent parts using dynamic programming \cite{wordninja} and lemmatize 
each individual sub-token. 
For example the method \texttt{\_\_zone\_seqlock\_init} corresponds to \texttt{zone}, \texttt{seqlock} and \texttt{init}, 
\texttt{inprogress} is split into \texttt{in} and \texttt{progress} and \texttt{literals} becomes \texttt{literal}.

\paragraph{Call Graph Generation} 
The codebase is processed by a Static Graph Analyzer. 
In our approach, focusing on C projects, we use \emph{CScout} \cite{cscout} in order to generate the directed call graph through function calls between the files. 
Each source code file is assigned to a module. 
Modules can be user defined or automatically generated by their respective directories at a desired directory tree depth.

@@

\paragraph{Experimental Setting} Our method was run on Linux 4.21,
consisting of 20.3 million SLOC against Average-Linkage \cite{average}, Complete-Linkage \cite{complete} 
and Ward-Linkage \cite{ward} using the same document embeddings
as well as ACDC with structural information \cite{acdc} and LIMBO \cite{limbo} with $B = 100$, $S = \infty$ and Bag-of-Words features. The results are presented in Table \ref{tab:evaluation}. 
As ground truth we have used the first level directories as a target clustering and as input we have considered 
the modules of the one-top directories. 
For example, the source code file \texttt{drivers/net/ieee802154\-/mcr20a.c} has a ground truth value of \texttt{drivers} 
and it is considered under the same module as every \texttt{.c} and \texttt{.h} file under \texttt{drivers/net/ieee802154}. 
As a clustering distance measure we have used the MoJo distance measure, as proposed in reference \cite{mojo}. 
Since, Louvain Community Detection produces  different (but very similar) results every time it runs, we ran the simulations 
multiple times and averaged the results.

@@

In Table \ref{tab:evaluation}, we present the results of the clusterings ordered in increasing MoJo distance 
with respect to the Ground Truth. 
We have included the number of clusters produced, the cluster size range, the average cluster size the standard
deviation in the cluster sizes, the median cluster and the MoJo distances with respect to the ground truth. 
The Ground Truth clustering is also present in the table. 
Our quality metrics for the clusterings are 
\emph{extremity}, \emph{authoritativeness} and \emph{stability} as proposed in reference \cite{evaluation}.

Firstly, our approach produces balanced clusterings thus demonstrating low \emph{extremity} compared to the other clusterings, 
producing a number of clusters that was close to the Ground Truth with smaller standard deviation in the cluster sizes than 
the other clustering methods, without knowing the number of clusters \emph{a priori}. 
Besides this, our approach also gives balanced clusters with respect to \emph{Median Cluster} criterion. 
Note that ACDC, which takes into account only the structural properties of the files, produced a high 
granularity clustering that failed to catch the structure of Linux Codebase. 
Moreover, in the \emph{authoritativeness} criterion, our approach outperforms the other clusterings to a very up to 10 times 
in terms of MoJo distance being very close (smaller is better) to the Ground Truth, with a very small standard deviation. 
This also suggests great \emph{stability} in the resulting clusterings. 
Note that the other clustering methods we compared against ours gave disappointing MoJo distances even with their cluster
number defined equal to the one of the Ground Truth.

@@

Onaiza Maqbool and Haroon A. Babri \cite{maqbool_overview} present an
overview of the various approaches taken towards to hierarchical clustering algorithms
for software systems. In their study, they compare the various clustering algorithms
such as ACDC, LIMBO and other Traditional Clustering Algorithms such as Single
Linkage, Complete Linkage and Weighted Linkage presenting their efficiency regarding 
multiple feature sets. 
Added to this, a \emph{semantics-based architectural view} of the system, as discussed in reference
\cite{large_study} reveals significant aspects of a software system and its change over 
time, which suggests that semantic-based approached should be followed for gaining better 
understandings of a software system. Their suggestion in \cite{large_study} is a \emph{term frequency--inverse 
document frequency} (TF-IDF) model serving as a similarity measure between source code 
entities. 
Work in Software Clustering Algorithms (SCA) has been extensively done by Tzerpos in
references \cite{mojo}, \cite{stability}, \cite{acdc} and \cite{limbo} with the ACDC
and LIMBO clustering algorithms.

In particular, the ACDC algorithm leans toward to software components comprehension 
based on subsystem patterns. Their approach considers an initial structure of the system, 
without taking into account semantics, and tries to build comprehensive clusterings of the given ground structure. 
The authors also conducted a study on an older version of  Linux. 
Moreover, Andritsos and Tzerpos in reference \cite{limbo} present an Information-Theoretic 
approach of SCA by developing LIMBO which clusters modules upon inserting their Distributional Cluster Features 
to a B+-tree variant and then applying the \emph{Agglomerative Information Bottleneck} algorithm, merging the 
clusters with the minimum mutual information loss together.

\section{Conclusions} This paper is set out to show how we can use a combination of vector semantics
and information from the call graph in order to produce meaningful clusterings of a software system.
We conclude that we have outperformed state-of-the-art software clusterings and conventional
agglomerative clustering algorithms 
in terms of \emph{extremity}, \emph{authoritativeness} and \emph{stability} 
without knowing the number of clusters of the Ground Truth. This evidence strongly
supports the further usage of vector semantics and the call graph for architecture recovery.
Finally, further work is encouraged in several areas such as 
integration with more static analyzers and use of the project in order to recover layered architectures.
